---
title: "Question 1"
author: "Name: swatee, simran, david"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

In the green buildings assignment, we are trying to study if green rating of a building has any effect on the rent it generates. And, if it does have effect then how much more or less rent does a green rated building generate compared to a building with similar features which is not green rated. The features of a building that we are studying along with green rating are cluster rent, size, employment rate in neighborhood, occupancy of the building, number of stories, age, renovated or not, quality of building (class a and class b), rent charged net of utility, closeness to shopping mall, bank, etc. heating and cooling in building, rain in the region, gas cost, and electricity cost. Studying these features will help us control for effects that these features have on the building rent. Hence, we will be able to isolate the effect of green rating on a building’s rent.

## Data and Model:

Firstly, we looked at the distribution of rent and cluster rent and found both data sets to be right skewed. So, we decided to transform rent and cluster_rent using log functions.
Then, we found that rent and cluster rent are highly correlated. We found a correlation coefficient of 0.7594093. This makes sense in real world. For instance, if a building is posh area where rent is high then most likely it will also have high rent. Hence, we thought it would be apt to standardize rent based on cluster_rent. Thus, we created a variable called y that stores standardized log rent.
Simultaneously, we created a sparse matrix of all the factor variables except for CS_PropertyID, LEED, Energystar, cluster_rent, cluster. We excluded cluster and cluster_rent because they are part of the dependent variable as we have standardized rent based on cluster rent. Then, we excluded CS_PropertyID because it is an identifier of the buildings and has no economic meaning. Finally, we excluded LEED and Energystar because they are the 2 types of green ratings that a building receives. Hence, including them would cause multicollinearity with green rating variable which is included our sparse matrix.
Afterwards, we fit a lasso regression in the data set and conduct cross validation with 10 folds.

```{r}
# import data and examine it
library(gamlr) 
greenbuildings <- read.csv("C:/Users/swate/Google Drive/MA Econ/spring 2020 classes/data mining/hw 3/greenbuildings.csv")
View(greenbuildings)
ok <- complete.cases(greenbuildings)
greenbuildings <- greenbuildings[ok,]

# check the correlation between rent and cluster rent
cor(greenbuildings$Rent, greenbuildings$cluster_rent)

# note that rent and cluster_rent are hugely skewed
# probably want a log transformation here
hist(greenbuildings$Rent)
summary(greenbuildings$Rent)

hist(greenbuildings$cluster_rent)

# much nicer :-)
hist(log(greenbuildings$Rent))
hist(log(greenbuildings$cluster_rent))

# i create a matrix of all my independent varaibles except for - CS_PropertyID - LEED -Energystar -cluster_rent -cluster  to make it easily readable for gamlr commands.
x = sparse.model.matrix( log(Rent) ~  . - CS_PropertyID - LEED -Energystar -cluster_rent -cluster , data=greenbuildings)[,-1] 

# I create a variable y that normalizes Log(rent) over log(cluster_rent)
y = (log(greenbuildings$Rent)- log(greenbuildings$cluster_rent))/sd(log(greenbuildings$cluster_rent))

# Here I fit my lasso regression to the data and do my cross validation of k=10 n folds
# the cv.gamlr command does both things at once.
cvl = cv.gamlr(x, y, nfold=10, verb=TRUE)
```

## Results:

On fitting the lasso regression we see that green rating coefficient is 0. Thus, green rating does not have any effect on the rent of the building.
We also plot lambda values and calculate the lambda value to see what penalty the lasso regression puts while minimizing the mean squared error. Lamda is -6.191394 and slopes of 16 variables are used in the penalty.

```{r}
# plot the out-of-sample deviance as a function of log lambda
plot(cvl, bty="n")

## CV min deviance selection
b.min = coef(cvl, select="min")
log(cvl$lambda.min) # this gives the value of lamda
sum(b.min!=0) # this gives the coefficent not 0

coef(cvl)
```

## Conclusion:

Therefore, we conclude that while size, age, class_a, amenities, and Gas_costs have significant impact on rent green rating does not effect rent. In real world, this seems plausible as most people would care about the size of the space, its quality indicated by age and class_a, its locality indicated by amenities, and gas costs when choosing a place and deciding on the rent rather than if it’s a green rated building or not.

